Cascades, Product Conversion, Whatever

### 2012-12-12 paper.tex

The analysis of strategic behavior on networks is not a new concept. Kearns et. al.
introduced a convenient method for expressing games of this sort with their seminal
work on graphical games \cite{Kearns01}. But, to the extent of our knowledge,
nobody has attempted to characterize full graphical game models for cascade spreading.
This is what we to below.

\section{Problem Statement}
\label{prob_statement}
We propose studying a simple cascade model over an arbitrary social
network. In this model each agent is represented by a node in a graph
and all of their friends / neighbors are represented by edges in the
same graph. Over the course of the game each agent will make an
irreversible choice between one of two types ($Y$ or $N$) -- where
agents that prefer $Y$ are less frequent -- and will get a payoff at
the end of the game corresponding to the number of neighbors that
chose the same type as well as a bonus if they chose their prefered
type. Thus each agent is torn between choosing their prefered type and
choosing the type that a majority of their neighbors chose. The game
is played with a single node making a choice between types at each
point in time. We are looking to choose an order or schedule over
agents to maximize the number of minority ($Y$) choices. We are
interested in two different classes of schedules, offline and
online. Offline schedules must schedule the entire set of agents
before the game is played, whereas online schedules can be revised as
agents make their decisions. We are interested in the difference
between the optimal number of minority types chosen when the agents
behave myopically, meaning their choice optimizes their current
payoff, and when the agents behave strategically, meaning their choice
optimizes their expected payoff at the end of the game. Strategic
agents vary significantly depending on how much information they
have. Here we assume they know the entire graph structure, every other
agents strategy, the probability of an agent’s type, all decided
node's choices, and the schedule.

Formally the model is specified for a graph $G(V,E)$ where each agent
is a node $v_i \in V, i={1..n}$. At every point in time $t={0..n}$
each agent is in one of three states, Yes, No, or Undecided ($v_{it}
\in \{Y, N, U\}$). At time zero each agent starts off Undecided, then
at each point in time one agent according to the schedule switches
from Undecided to either Yes or No. The schedule is an ordered
permutation ($\mathcal P$) of the nodes. For the offline schedule this
is static a permutation chosen ahead of time, but in the online
schedule the $j^{th}$ element of the permutation is a function of the
graph after the previous time. If the permutation is definied as an
$n$ vector of ${1..n}$ then $\mathcal P_t = f_{\mathcal M}(V_t)$ where
$\mathcal P_t$ is the $t^{th}$ element of the permutation,
$f_{\mathcal M}$ is an arbitrary function that specifies which node to
choose next based off of the current graph and the model parameters
$\mathcal M$, and $V_t$ is the state of the agents at time $t$. The
optimal schedule is the schedule that maximizes $J = \sum_{i=1}^n
\mathbb E[\mathbb I(v_{in} = Y)]$ with respect to the schedule
$\mathcal P$, where $\mathbb I$ is the indicator function and $v_{it}$
is the state of agent $i$ at time $t$. Before the game is played each
agent also gets a prefered type which is drawn from a bionmial where
$p(type_i = Y) = p < \frac{1}{2}$. The utility of a node $i$ at time
$t$ is $u_{it}(v_{it}) = \pi \mathbb I(v_{it} = type_i) + \sum_{j \in
  \mathcal N(i)} \mathbb I(v_{it} = v_{jt}), v_{it} \in \{Y, N\}$,
where $\pi$ is a constant that determines how much utility an agent
gets from choosing its prefered type. If an agent is deciding its type
at time $t$ a myopic agent $i$ will choose to maxmimize its current
utility $\arg\max_{v_{it} \in \{Y, N\}} u_{it}$ where as a strategic
agent will choose to maximize its expected final utility
$\arg\max_{v_{it} \in \{Y, N\}} \mathbb E[u_{in}]$. We are interested
in looking at the difference between the optimal number of minority
choosing agents when the agents are strategic versus myopic
($J_{myopic} / J_{strategic}$).

Ongoing work by Martin et al. suggests that the scheduler payoff is
always lower in the strategic case than in the myopic case, but we are
interested quantifying this difference. To tackle this problem we
propose investigating efficient algorithms for finding the optimal
scheduling strategies on specific classes of graphs, which we hope
will provide insight into quantifying the difference between the
expected number of agents picking the minority type when the agents
are strategic versus myopic.

We believe that an efficient dynamic programming algorithm exists for
finding the optimal schedule on graphs that can be specified as a
$2$-block graph. We define a $k$-block graph as any graph whose
adjacency matrix can be represented as symmetric block matrix with
$k^2$ blocks such that each block is either all $1$'s or all
$0$'s. The idea behind the algorithm is to do a pruned backward
induction from potential termination states. Because agents in a
single block are substitutable, the potential choice at each round is
significantly reduced. This combined with pruning should provide the
necessary requirements to make this algorithm efficient. Work by
Chierichetti et al.\cite{Chierichetti12} has already bounded the
optimal schedule payout for arbitrary graphs, however applying an
algorithm similar to ours for strategic agents should be no worse in
complexity for myopic agents. In addition to the optimal offline
schedule, the inductive nature of this algorithm should mean it also
calculates the optimal online schedule. This efficient optimal
algorithm should allow us to calculate the difference between
strategic and myopic agents under offline and online schedules, and
will hopefully give insight into bounds on the difference for
arbitrary 2-block graphs.

\section{Model}

Model Description goes here... This probably should steal a lot of
problem statement. My thoughts are this should define the game

\subsection{Block Models}

Then maybe a subsection specifically about block models

\begin{align*}
  & k && \text{Number of blocks} \\
  & \{s_i\}_{i=1..k} && \text{Number of nodes in each block}
\end{align*}

@@

The algorithm uses two lookup tables (one for the scheduler, and one
for the nodes) to determine the optimal play given the relevant
statistics of the game at every decision time. Each table has a
dimension for every relevant play statistic for the deciding agent,
and contains their optimal choice as well as the expected number of
Yeses in each block after that choice is made (which is necessary for
callers to decide on their optimal decision, as expected payoffs are
only a function of the expected number of Yeses at the end of the
game).

@@

>>>>

The analysis of strategic behavior on networks is not a new
concept. Kearns et. al.  introduced a convenient method for expressing
games of this sort with their seminal work on graphical games
\cite{Kearns01}. But, to the extent of our knowledge, nobody has
attempted to characterize full graphical game models for cascade
spreading.  This is what we do below.

\section{Model}
\label{prob_statement}

We propose a cascade spreading model that borrows heavily from the
scheduler mediated cascade process introduced in
\cite{Chierichetti12}. In our model every node in an undirected graph
is a risk neutral rational agent. Each node is also assigned a binary
type which affects their utility, pulled from a binomial
distribution. In this paper we call the types Yes ($Y$) and No
($N$). There is an additional scheduler agent, which, at each round of
the game picks one of the nodes in the graph. At that moment the
chosen node chooses to commit to a choice (also Yes or No) and the
agent is committed to that choice for the rest of the game. The game
ends when all nodes have committed to a choice.

We keep the same scheduler behavior as in the previous paper. The
schedulers utility is based off of the number of nodes that choose Yes
at the end of the game. We assume the scheduler can change its plan
after it sees what a node chooses. We also assume it has full
knowledge of the nodes strategies, as well as knowledge of the full
graph, and any game parameters. The only aspect the scheduler does not
know is the individual nodes types.

The nodes utility is the sum of all of their neighbors that match
their choice plus some constant ($\pi$) if their choice matches their
type. Strategic nodes make decisions to maximize their expected
profit. We assume strategic nodes know the schedulers decision making
(or utility function), the graph structure, and the probability that a
node is assigned Yes type ($p$), but not any other nodes type. To keep
the model interesting, we restrict $p < 0.5$. Similarly myopic nodes
only try to maximize their current utility, ignoring any nodes that
has not made a decision.

We consider games where either all nodes in the graph are strategic or
myopic, and want to see what graphs / game parameters contribute to
significant differences between the expected number of Yesses when
nodes are strategic versus myopic.

We formally specify the model as follows
\begin{align*}
  & G(V,E) && \text{social network} \\
  & |V| = n \\
  & c_i \in \{Y,N\} && \text{choice of node $i$} \\
  & t_i \in \{Y, N\} && \text{type of node $i$} \\
  & p(t_i = Y) = p < 0.5 \\
  & u_i(\mathbf c) = \pi \mathbb I[t_i = c_i] + \sum_{j \in N(i)}
  \mathbb I[c_j = c_i] && \text{node utility} \\
  & sched(\mathbf c) && \text{schedule function}
\end{align*}
where $\mathbb I$ is the indicator function, and $N$ is the neighbor
function. The schedule function returns the next node to pick given
the choices that have been made so far.

The full game can be described by the following algorithm. The
scheduler picks an optimal first node in the graph. That node chooses
either Yes or No based on all of the parameters and its
knowledge. Based on this result the scheduler picks a new node, and so
on until every node is chosen and the game finishes.

Given this game, we are interested in quantifying
\begin{equation*}
  \frac{\mathbb E[\#Y|strategic]}{\mathbb E[\#Y|myopic]}.
\end{equation*}

Is seems likely that computing the strategic choices and payouts is a
hard problem, but we will show an algorithm that can compute an
optimal online schedule and calculate $\mathbb E[\#Y|stragetic]$ in
polynomial time for arbitrary block models.

A graph is a $k$-block model is the vertices can be divided into $k$
disjoint subsets such that every node in the same subset has
incoming and outgoing edges to the same set of nodes. Another way to
view this property is by looking at the adjacency matrix of the
graph. A block model graph can be written as a symmetric block matrix
with $k^2$ blocks. For the sake of the algorithm, we will define
\begin{equation*}
  \{s_i\}_{i=1..k}
\end{equation*}
as the number of nodes in each block.

@@

Our algorithm uses two lookup tables (one for the scheduler, and one
for the nodes) to determine the optimal play given the relevant
statistics of the game at every decision time. Each table has a
dimension for every relevant play statistic for the deciding agent,
and contains their optimal choice as well as the expected number of
Yeses in each block after that choice is made (which is necessary for
callers to decide on their optimal decision, as expected payoffs are
only a function of the expected number of Yeses at the end of the
game).

@@

Patterns may not exist, but there may still be order level
relationships / bounds that are possible.

See if we can prove the problem is PPAD or NP-complete.

Offline scheduling
